---
title: "Fallopia Analysis"
output: html_document
date: "2023-02-23"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note this Rmd is the abbreviated version of "Nallocation_Fallopia.Rd" that has all the method details (esp. on N partitioning).

## Dataset

The dataset (dat) is the merged output of the Bayesian photosynthesis model (from the script 'photo_params_all_HB5.R', including posterior means and SDs of Asat (A at 40 Pa and saturating light), Vcmax, Jmax, and Rd) and lab assay results in the master Google Docs spreadsheet (currently version 7), plus a separate csv file of associated covariates for each sample (nativity, range, growth form).

```{r}
dfold = "C:\\Users\\fridley\\OneDrive - Clemson University\\academic\\projects\\IOS_FranceJapan\\"
dat = read.csv(paste0(dfold,"FAJA_data_final3.csv")) #v2 includes new cyanogenic glycoside N values
  #v3 includes updated soil CN, pH
dat$region = as.factor(substr((dat$Region),1,1))
covar = read.csv(paste0(dfold,"covariates.csv")) #covariates at the species-region level
dat = merge(dat,covar,by.x=c("Species","region"),by.y=c("species","region"),all.x=T)
temp.dat = read.csv(paste0(dfold,"leaftempsbyID.csv")) #file of median leaf temps per sample
names(temp.dat) = c("ID","Tleaf")
dat = merge(dat,temp.dat,by.x="ID",by.y="ID",all.x=T) #merge leaf temps with dataset
```

### Colors for plotting

```{r}
#colors    
  makeTransparent<-function(someColor, alpha=150)
  {   newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3],alpha=alpha, maxColorValue=255)}) }
c1 = makeTransparent("darkgreen") #ENA
c2 = makeTransparent("purple")  #France
c3 = makeTransparent("darkorange") #Japan
colvec = c("darkgreen","purple","darkorange")
```

## Soils Data

We have pH, soil N, and soil C. May be most efficient to examine pH and C:N ratios.

```{r}
library(RColorBrewer)
bar.cols = brewer.pal(n = 8, name = 'Greens')[c(8,7,5,1)]
bar.cols = c(bar.cols[1:3],"burlywood4","orange","darkgray")

#pH
summary(dat$soil_pH) #on average about 10 mg per g leaf, this is low
boxplot(dat$soil_pH~dat$Region,col=c(c1,c2,c3)) 
summary(aov(dat$soil_pH~dat$Region)) #Japan nearly 2 units pH lower, P<0.001

#soil CN
summary(dat$soil_CN) #on average about 10 mg per g leaf, this is low
boxplot(dat$soil_CN~dat$Region,col=c(c1,c2,c3)) 
summary(aov(dat$soil_CN~dat$Region)) #no significant difference in soil CN across regions, P=0.075

```

## Leaf N pools

We are interested in 5 pools of leaf N:
1. Nr = N in Rubisco. Measured assuming protein is 16% N.
2. Nl = N allocated to light harvesting, calculated according to Takashima et al. (2004).
3. Nb = N allocated to bioenergetics associated with electron transfer proteins, according to Takashima et al. (2004).
4. Nc = N associated with cell wall proteins, measured directly.
5. Nd = N associated with defensive secondary metabolites, esp. cyanogenic glycosides and alkaloids, measured by Bensaddek in France with N assumed as fraction according to most common compounds.


```{r warning=F, fig.width=8, fig.height=4}

#Rubisco N (Nr)
dat$SLA = as.numeric(as.character(dat$Specific.leaf.area)) #SLA in cm2 g-1
dat$Rub.g = as.numeric((dat$Rubisco.content)) * dat$SLA #ug Rubisco per g leaf
dat$RubN = dat$Rub.g * .16 #ug N in Rubisco per g leaf
dat$Nmass = (as.numeric(as.character(dat$leafN))/100) #g leaf N per g leaf
dat$Rub.perc = (dat$RubN / 1000000) / dat$Nmass #Rubisco g per g leaf N

#Light harvesting N (Nl)
dat$chl.ugcm2 = as.numeric(as.character(dat$Total.Chlorophyll)) #ug chl per cm2
dat$chl.mmolm2 = (dat$chl.ugcm2 / 1000000) * 1000 * 10000 / 893.5 #mmol chl per m2 leaf
dat$chl.ugg = dat$chl.ugcm2 * dat$SLA  #ug chl per g leaf
dat$chlN = dat$chl.ugg / (1/.0627) #ug N in Chl per g leaf
dat$Chl.perc = (dat$chlN / 1000000) / dat$Nmass
dat$chl.mmolg = dat$chl.mmolm2*dat$SLA/10000 #mmol chl per g leaf
dat$Nl.NT = dat$chl.mmolg / (dat$Nmass*1.93) #N&T '97 approach, assuming Cb is 1.93
dat$Nl.Taka = (dat$chl.mmolg * 37.1 / 1000) * 14 * (1/dat$Nmass) #Takashima method, similar

#Bioenergetics N (Nb)
Jmc.temp = function(c,dHa,dHd,dS,R,Tk) { #temperature dependent
  exp(c-(dHa/(R*Tk))) / (1 + exp(  ((dS*Tk)-dHd) / (R*Tk)) )
}
Tk.vec = seq(20,40,length=100) + 273.15
dat$Tleaf[is.na(dat$Tleaf)] = 30
dat$Jmc = Jmc.temp(c=14.77, dHa=24100, dHd=564150, dS=1810, R=8.314, Tk=dat$Tleaf+273.15)
dat$Nb.Taka = dat$Jmax*(9.53/dat$Jmc)*(dat$SLA/10000)*14*(1/dat$Nmass)/1000
dat$Nb.NT = dat$Jmax * (dat$SLA/10000) * (1/dat$Nmass) * (1/(8.06*dat$Jmc)) #similar

#Cell wall N (Nc)
dat$wallN = dat$Cell.wall..N / 100 #g N per g cell wall
dat$Narea.cm2 = dat$Nmass / dat$SLA #g N per cm2 leaf area
dat$wallg.percm2 = (dat$Cell.wall.sample.mass.mg / as.numeric(as.character(dat$cell.wall.leaf.area.cm2)))/1000 #cell wall mass, g per cm2 leaf area
dat$wallg.perg = dat$wallg.percm2*dat$SLA #cell wall mass as % leaf mass
dat$Nc = (dat$wallg.percm2 * dat$wallN) / (dat$Narea.cm2) #percent leaf N present in cell wall

#Defensive N (Nd)
dat$defenseN = dat$total_alkaloidN_mg_g/1000 + dat$cyanN_mg_g/1000 #sum of alkaloids and cyanogenic glycosides
```

Leaf phenolics are also involved in defense but don't contain N. But the enzymes required to synthetic phenolics do of course contain N, so they may be relevant:

```{r}
summary(dat$total_polyphenols_mg_g) #on average about 10 mg per g leaf, this is low
boxplot(dat$total_polyphenols_mg_g~dat$Region,col=c(c1,c2,c3)) #no differences across regions
```

Summary of leaf N fractions in Fallopia across regions:

```{r}

dat$Narea = dat$Nmass / (dat$SLA/10000)
dat$NUE = dat$Asat / dat$Narea
dat$Nthy = dat$Nl.Taka + dat$Nb.Taka
fall = dat
fall$Region[fall$Region=="ENA"] = "USA"
fall$Region[fall$Region=="Japan"] = "aJapan" #to have home range presented first

library(RColorBrewer)
bar.cols = brewer.pal(n = 8, name = 'Greens')[c(8,7,5,1)]
bar.cols = c(bar.cols[1:3],"burlywood4","orange","darkgray")

#Stacked bar graph based directly on means of partitions
FNr.m = tapply(fall$Rub.perc,fall$Region,function(x)mean(x,na.rm=T))
FNl.m = tapply(fall$Nl.Taka,fall$Region,function(x)mean(x,na.rm=T))
FNb.m = tapply(fall$Nb.Taka,fall$Region,function(x)mean(x,na.rm=T))
FNc.m = tapply(fall$Nc,fall$Region,function(x)mean(x,na.rm=T))
FNd.m = tapply(fall$defenseN,fall$Region,function(x)mean(x,na.rm=T))
Ntot.m = tapply(fall$Nmass,fall$Region,function(x)mean(x,na.rm=T))
barht.rel = rbind(FNr.m,FNl.m,FNb.m,FNc.m,FNd.m)
barht = barht.rel
barht[,1] = barht.rel[,1] * Ntot.m[1]
barht[,2] = barht.rel[,2] * Ntot.m[2]
barht[,3] = barht.rel[,3] * Ntot.m[3]
tot3 = colSums(barht)
FN.miss = Ntot.m - tot3 
barht2 = rbind(barht,FN.miss)
midpoints = apply(barht2,2,cumsum) - barht2/2  #height of middle of each component for plotting
midpoints[5,] = midpoints[5,] + .0005 #adjust defense because so small
percents = barht.rel*100; lastcol = 100-colSums(barht.rel)*100; percents = rbind(percents,lastcol)
perc.lab = paste0(round(percents,1),"%")
par(mar=c(3,5,1,6.5),mfrow=c(1,1))
b1 = barplot(barht2*1000,beside=F,ylab="mg N per N leaf",col=bar.cols,ylim=c(0,30),names.arg=c("Japan","France","USA"),
            cex.lab=1.5,cex.axis=1.4,cex.names=1.5)
#title(main="Fallopia N allocation per g leaf N")
text(rep(b1,each=6),as.numeric(midpoints)*1000,perc.lab,col="white",cex=1.4,lwd=2)
mtext("Rubisco",side=4,las=1,at=midpoints[1,3]*1000,cex=1.2)
mtext("Light harvest",side=4,las=1,at=midpoints[2,3]*1000,cex=1.2)
mtext("Bioenergetics",side=4,las=1,at=midpoints[3,3]*1000,cex=1.2)
mtext("Cell wall",side=4,las=1,at=midpoints[4,3]*1000,cex=1.2)
mtext("Defense",side=4,las=1,at=midpoints[5,3]*1000,cex=1.2)
mtext("Residual",side=4,las=1,at=midpoints[6,3]*1000,cex=1.2)
text(b1,c(20.8-.4,22-.4,26-.4),round(tapply(fall$leafN,fall$Region,mean),2),cex=1.4)

#same figure for pdf

pdf("Figure1.pdf",width=8,height=8)
par(mar=c(3,5,1,6.5),mfrow=c(1,1))
b1 = barplot(barht2*1000,beside=F,ylab="mg N per N leaf",col=bar.cols,ylim=c(0,30),names.arg=c("Japan","France","USA"),
            cex.lab=1.5,cex.axis=1.4,cex.names=1.5)
#title(main="Fallopia N allocation per g leaf N")
text(rep(b1,each=6),as.numeric(midpoints)*1000,perc.lab,col="white",cex=1.4,lwd=2)
mtext("Rubisco",side=4,las=1,at=midpoints[1,3]*1000,cex=1.2)
mtext("Light harvest",side=4,las=1,at=midpoints[2,3]*1000,cex=1.2)
mtext("Bioenergetics",side=4,las=1,at=midpoints[3,3]*1000,cex=1.2)
mtext("Cell wall",side=4,las=1,at=midpoints[4,3]*1000,cex=1.2)
mtext("Defense",side=4,las=1,at=midpoints[5,3]*1000,cex=1.2)
mtext("Residual",side=4,las=1,at=midpoints[6,3]*1000,cex=1.2)
text(b1,c(20.8-.4,22-.4,26-.4),round(tapply(fall$leafN,fall$Region,mean),2),cex=1.4)
dev.off()


```

The principal difference is the ENA populations have higher total N from higher Rubisco. France to a lesser extent. What allocation differences are statistically significant?

```{r}
man1 = manova(cbind(Rub.perc,Nl.Taka,Nb.Taka,Nc,defenseN)~Region,data=fall)
summary(man1)
  #relative allocation is significantly different
summary.aov(man1)
  #there are significant different in all allocation parameters if uncorrected
  #with corrected P values, rubisco, light harvesting, bioenergetics are different
    #cell wall and defense are not

#post hoc tests
TukeyHSD(aov(Rub.perc~Region,data=fall))
  #ENA and France greater than Japan

TukeyHSD(aov(Nl.Taka~Region,data=fall))
  #weak suggestion that France is higher than ENA and Japan

TukeyHSD(aov(Nb.Taka~Region,data=fall))
  #Japan and France much greater than ENA

TukeyHSD(aov(Nc~Region,data=fall))
  #no differences

TukeyHSD(aov(defenseN~Region,data=fall))
  #oddly, US higher than both japan and France


```

### Differences in photosynthetic parameters across regions

```{r}
boxplot(Asat~Region,data=dat,col=colvec)
boxplot(Vcmax~Region,data=dat,col=colvec)
boxplot(Jmax~Region,data=dat,col=colvec)
boxplot(Rd~Region,data=dat,col=colvec) #?? Rd much higher in France

summary(aov(Asat~Region,weights=1/Asat.se,data=dat))
summary(aov(Vcmax~Region,weights=1/Asat.se,data=dat))
summary(aov(Jmax~Region,weights=1/Asat.se,data=dat))
summary(aov(Rd~Region,weights=1/Asat.se,data=dat)) #P<0.001

dat$PNUE = dat$Asat/dat$Narea
boxplot(PNUE~Region,data=dat,col=colvec)
TukeyHSD(aov(PNUE~Region,data=dat)) #ENA smaller than others, P<0.01 for both contrasts 

dat$Amass = dat$Asat*dat$SLA/10000
boxplot(Amass~Region,data=dat,col=colvec)
summary(aov(Amass~Region,data=dat)) #ND

dat$Rdmass = dat$Rd*dat$SLA/10000
boxplot(Rdmass~Region,data=dat,col=colvec)
TukeyHSD(aov(Rdmass~Region,data=dat)) #France substantially higher Rd (double)

#figure for pub
pdf("Figure2.pdf",width=8,height=8)
par(mfrow=c(2,2),mar=c(5,5,1,1))
plot(dat$Narea,dat$Asat,col=c(c1,c2,c3)[as.numeric(as.factor(dat$Region))],pch=19,cex=2,
    xlab=expression('N'[area]*' (g m'^-2*')'),cex.lab=1.5,cex.axis=1.5,
    ylab=expression('A'[sat]*' (' * mu * 'mol CO'[2]*' m'^-2*'s'^-1*')'))
points(dat$Narea,dat$Asat,col=c(colvec)[as.numeric(as.factor(dat$Region))],pch=21,cex=2,lwd=2)
summary(lm(Asat~Narea,data=dat,subset=dat$Region=="Japan")) #P=0.1
abline(lsfit(dat$Narea[dat$Region=="Japan"],dat$Asat[dat$Region=="Japan"]),col=colvec[3],lwd=3,lty=1)
summary(lm(Asat~Narea,data=dat,subset=dat$Region=="ENA")) #P=0.02
abline(lsfit(dat$Narea[dat$Region=="ENA"],dat$Asat[dat$Region=="ENA"]),col=colvec[1],lwd=3,lty=1)
summary(lm(Asat~Narea,data=dat,subset=dat$Region=="France")) #P=0.03
abline(lsfit(dat$Narea[dat$Region=="France"],dat$Asat[dat$Region=="France"]),col=colvec[2],lwd=3,lty=1)

par(mar=c(5,1,1,5))
boxplot(Asat~Region,data=fall,col=colvec[c(3,2,1)],cex.axis=1.5,cex.lab=1.5,ylab="",xlab="",xaxt="n",yaxt="n",names.arg=c("Japan","France","USA"))
axis(1,labels=c("Japan","France","USA"),at=c(1,2,3),cex.axis=1.5)
axis(2,labels=F)

par(mar=c(5,5,1,1))
plot(dat$Rubisco.content,dat$Vcmax,col=c(c1,c2,c3)[as.numeric(as.factor(dat$Region))],pch=19,cex=2,
    xlab=expression('Rubisco'*' (' * mu *  'g cm'^-2*')'),cex.lab=1.5,cex.axis=1.5,
    ylab=expression('V'[cmax]*' (' * mu * 'mol CO'[2]*' m'^-2*'s'^-1*')'))
points(dat$Rubisco.content,dat$Vcmax,col=c(colvec)[as.numeric(as.factor(dat$Region))],pch=21,cex=2,lwd=2)

summary(lm(Vcmax~Rubisco.content,data=dat,subset=dat$Region=="France")) #P=0.76
summary(lm(Vcmax~Rubisco.content,data=dat,subset=dat$Region=="Japan")) #P=0.56
summary(lm(Vcmax~Rubisco.content,data=dat,subset=dat$Region=="ENA")) #P=0.90


par(mar=c(5,1,1,5))
boxplot(Vcmax~Region,data=fall,col=colvec[c(3,2,1)],cex.axis=1.5,cex.lab=1.5,ylab="",xlab="",yaxt="n",xaxt="n")
axis(2,labels=F)
axis(1,labels=c("Japan","France","USA"),at=c(1,2,3),cex.axis=1.5)
dev.off()

```

### Differences in other leaf traits across Regions

```{r}
boxplot(SLA~Region,data=dat,col=colvec)
TukeyHSD(aov(SLA~Region,data=dat)) #France higher than others, P<0.01 for both contrasts

boxplot(Narea~Region,data=dat,col=colvec)
TukeyHSD(aov(Narea~Region,data=dat)) #ENA higher than others, P<0.01 for both contrasts

boxplot(Nmass~Region,data=dat,col=colvec)
TukeyHSD(aov(Nmass~Region,data=dat)) #ENA higher than others, P<0.01 for Japan but not France (.06)

boxplot(leafC~Region,data=dat,col=colvec)
TukeyHSD(aov(leafC~Region,data=dat)) #reduction in leaf C in away range, sig (P=.01) for ENA

dat$Carea = dat$leafC / (dat$SLA/100)
boxplot(Carea~Region,data=dat,col=colvec)
TukeyHSD(aov(Carea~Region,data=dat)) #pairwise diff except ENA-Japan


boxplot(leaf.delta15N~Region,data=dat,col=colvec)
TukeyHSD(aov(leaf.delta15N~Region,data=dat)) #ENA and France higher than Japan, unclear meaning

boxplot(leafN.fall~Region,data=dat,col=colvec) 
TukeyHSD(aov(leafN.fall~Region,data=dat))
summary(aov(leafN.fall~Region,data=dat))
  #trend toward lower N in senesced/autumn ENA populations 

dat$CN = dat$leafC / dat$leafN
boxplot(CN~Region,data=dat,col=colvec) 
TukeyHSD(aov(CN~Region,data=dat))

dat$resorbNeff = dat$leafN.fall / dat$leafN
boxplot(resorbNeff~Region,data=dat,col=colvec) 
summary(aov(resorbNeff~Region,data=dat)) #trend toward higher N resorption efficiency in Japan

dat$chlab.ratio = dat$Chlorophyll.a / dat$Chlorophyll.b
boxplot(chlab.ratio~Region,data=dat,col=colvec) 
TukeyHSD(aov(chlab.ratio~Region,data=dat)) #significantly higher in Japan

#Fiber
man.fiber = manova(cbind(dat$Fiber..solubles...NSCs.,dat$Fiber..hemicellulose.,dat$Fiber..cellulose,dat$Fiber..lignin.,dat$Fiber..mineral.content.)~Region,data=dat)
summary(man.fiber)  
summary.aov(man.fiber)
  #very large differences in hemicellulose and cellulose

boxplot(Fiber..solubles...NSCs.~Region,data=dat,col=colvec) 
TukeyHSD(aov(Fiber..hemicellulose.~Region,data=dat)) #striking differences across regions, ENA then France then Japan


boxplot(Fiber..hemicellulose.~Region,data=dat,col=colvec) 
TukeyHSD(aov(Fiber..hemicellulose.~Region,data=dat)) #striking differences across regions, ENA then France then Japan

boxplot(Fiber..cellulose.~Region,data=dat,col=colvec) 
TukeyHSD(aov(Fiber..cellulose.~Region,data=dat)) #same story with cellulose, ENA then France then Japan

#check soil pH
boxplot(Soil.pH~Region,data=dat,col=colvec) 
TukeyHSD(aov(Soil.pH~Region,data=dat))
  #Japan soils much more acidic, may impact N availability
  #but ENA and France also differed in leaf N despite no pH difference
  #soil CN data look like a lost cause


```

### Calculation of iWUE (integrated Water Use Efficiency)

```{r}
#C3 equation based on  Farquhar and Richards (1984) with -8 ppt ambient air
  # C triangle = (d13C.air - d13C.plant) / (1+(del13C.plant/1000))
dat$Cdiscrim = (-8 - dat$leaf.delta13C) / (1 + (dat$leaf.delta13C/1000))

#iWUE from Farquahar et al. 1989, assuming ambient CO2 conc of 400 ppm
  # iWUE = CO2ambient.ppm*(27 - Cdiscrim) / (1.6*(27-4.4))
  #**need to check this, could relate to instantaneous WUE (A/g) from licor files
dat$iWUE = 400 * (27-dat$Cdiscrim) / (1.6*(27-4.4))
boxplot(iWUE~Region,data=dat,col=colvec)
TukeyHSD(aov(iWUE~Region,data=dat)) #Japan much higher iWUE, but check whether this is related to altitude

#remains true for leaves sampled in fall
boxplot(leaf.delta13C.fall~Region,data=dat,col=colvec)
TukeyHSD(aov(leaf.delta13C.fall~Region,data=dat))

```

### Calculation of leaf construction costs.

```{r}
#from Heberling & Fridley 2013 in glucose equivalents
  #CCmass = (-1.041 + 5.077*Cmass)*(1-.67*Ash) + 5.325Nmass
dat$ash = dat$Fiber..mineral.content. * 100
  #negative values converted to zero
  dat$ash[dat$ash<0] = 0
dat$CCmass = (-1.041 + (5.077*(dat$leafC/100)))*(1-(.67*dat$ash)) + 5.325*(dat$leafN/100)

boxplot(dat$CCmass~Region,data=dat,col=colvec)
TukeyHSD(aov(CCmass~Region,data=dat))
  #marginally different between ENA and France

dat$CCarea = dat$CCmass*10000/dat$SLA
boxplot(dat$CCarea~Region,data=dat,col=colvec)
TukeyHSD(aov(CCarea~Region,data=dat)) #France is lower mainly due to high SLA


```

### Traits bar graph for pub

```{r}

pdf("Figure3.pdf",height=8,width=8)
par(mfrow=c(3,3),mar=c(2,5,1,1))
#Narea
x = dat$Narea
lab = expression(N[area]*' (g m'^-2*')')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(0,max(mse)),ylab=lab,cex.lab=1.5,cex.axis=1.3)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#SLA
x = dat$SLA
lab = expression('SLA (g cm'^-2*')')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(0,max(mse)),ylab=lab,cex.lab=1.5,cex.axis=1.3)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#chl
x = dat$chl.ugcm2
lab = expression('Chl (' * mu * 'g cm'^-2*')')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(0,max(mse)),ylab=lab,cex.lab=1.5,cex.axis=1.3)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#chl ab
#x = dat$Chlorophyll.a / dat$Chlorophyll.b
#lab = expression('Chl a:b')

#PNUE
dat$PNUE = dat$Asat/dat$Narea
x = dat$PNUE
lab = expression('PNUE ('*mu*'mol CO'[2]*' g'^-1*' N'*' s'^-1*')')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(0,max(mse)),ylab=lab,cex.lab=1.5,cex.axis=1.3)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#WUEi
x = dat$iWUE
lab = expression('WUE ('*mu*'mol CO'[2]*' mol'^-1*' H'[2]*'O)')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(0,max(mse)),ylab=lab,cex.lab=1.5,cex.axis=1.3)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#resorption
dat$resorbNeff = (dat$leafN-dat$leafN.fall) / dat$leafN
x = dat$resorbNeff
lab = "N resorption (%)"
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(0,max(mse)),ylab=lab,cex.lab=1.5,cex.axis=1.3)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#CCmass
x = dat$CCmass
lab = expression('CC'[mass]*' (eq. g glucose g'^-1*')')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(1,1.5),ylab=lab,cex.lab=1.5,cex.axis=1.3,xpd=F)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#Cmass
x = dat$leafC
lab = expression('C'[mass]*' (%)')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(44,48),ylab=lab,cex.lab=1.5,cex.axis=1.3,xpd=F)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)

#Fiber
x = dat$Fiber..hemicellulose. + dat$Fiber..cellulose
lab = expression('Hemi + Cellulose (%)')
m = tapply(x,fall$Region,function(x)mean(x,na.rm=T))
se = tapply(x,fall$Region,function(x)sd(x,na.rm=T)/sqrt(length(x[!is.na(x)])))
mse = (m + se)*1.3
b1 = barplot(m,names.arg=c("Japan","France","USA"),col=colvec[c(3,2,1)],xaxt="n",ylim=c(0,max(mse)),ylab=lab,cex.lab=1.5,cex.axis=1.3)
arrows(b1,m,b1,m+se,code=2,length=.1,angle=90,lwd=2)
dev.off()

```

### Relationship between leaf properties and photosynthetic parameters

```{r}
#Asat vs. Rubisco vs. Narea 
fall$Region = as.factor(fall$Region)
par(mfrow=c(2,2),mar=c(4,4,1,1))
plot(fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Leaf N (g m-2)",ylab="Asat")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c(c1,c2,c3),"bottomright")
plot(fall$Rub.perc*fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Rubisco N (g m-2)",ylab="Asat")
plot(fall$Nl.Taka*fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Light harvest N (g m-2)",ylab="Asat")
plot(fall$Nb.Taka*fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Bioenergetics N (g m-2)",ylab="Asat")
  #this is because Jmax drives Asat in this species at 40 Pa Ci and saturating light
  #plot(fall$Vcmax,fall$Asat)
summary.aov(lm(Asat~Narea*Region,data=fall))
TukeyHSD(aov(Asat~Narea*Region,data=fall),which="Region") #only weak evidence of difference

#Rubisco should better correlate with Vcmax
par(mfrow=c(1,2),mar=c(4,4,1,1))
plot(fall$Narea,fall$Vcmax,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Leaf N (g m-2)",ylab="Vcmax")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c(c1,c2,c3),"topleft")
plot(fall$Rub.perc*fall$Narea,fall$Vcmax,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Rubisco N (g m-2)",ylab="Vcmax")
  #but it doesn't; suggests a problem with Rubisco?
summary(lm(Vcmax~Rubisco.content,data=fall))
  #no relationships

```

I think this is the largest weakness of our dataset--there is no clear relationship between photosynthetic properties and leaf N allocation.


### Multivariate ordination and tests

```{r fig.width=8, fig.height=8}
library(pcaMethods)

#choose traits for inclusion
#tr.select = c("leafC","Cell.wall.mass","SLA","Total.Chlorophyll","Fiber..solubles...NSCs.","Fiber..hemicellulose.","Fiber..cellulose.","Fiber..lignin.","ash","total_alkaloids_mg_g","cyanN_mg_g","total_polyphenols_mg_g","Asat","Vcmax","Jmax","Rd","Rub.perc","Nmass","Narea","Nl.Taka","Nb.Taka","Nc","defenseN","PNUE","Amass","Rdmass","resorbNeff","chlab.ratio","iWUE","CCmass","CCarea")

#new order based on pub table
tr.select = c("Asat","Amass","Vcmax","SLA","Nmass","Narea","resorbNeff","leafC","Carea","CN","Total.Chlorophyll","chlab.ratio","CCmass","CCarea","iWUE","PNUE","Fiber..hemicellulose.","Fiber..cellulose.","Fiber..lignin.","ash","total_alkaloids_mg_g","cyanN_mg_g","total_polyphenols_mg_g","Rub.perc","Nl.Taka","Nb.Taka","Nc","defenseN")

traitmat = dat[,tr.select]
pcamat = scale(traitmat) #scale each  
ppca1 = pcaMethods::pca(pcamat,method="ppca",seed=123,nPcs=2)
ppca1 #PC 1 = 22%, PC 2 = 18%

#labels
pretty = names(pcamat)
pretty[8] = expression('C'[mass])
pretty[9] = expression('C'[area])
pretty[10] = expression("CN")
pretty[4] = expression("SLA")
pretty[11] = expression("Chl")
pretty[17] = expression("Hemi")
pretty[18] = expression("Cellulose")
pretty[19] = expression("Lignin")
pretty[20] = expression("Ash")
pretty[21] = expression("Alk")
pretty[22] = expression("Cyan")
pretty[23] = expression("Phenol")
pretty[1] = expression("Asat"[area])
pretty[3] = expression("V"[cmax])
pretty[24] = expression("N"[R])
pretty[5] = expression("N"[mass])
pretty[6] = expression("N"[area])
pretty[25] = expression("N"[L])
pretty[26] = expression("N"[B])
pretty[27] = expression("N"[C])
pretty[28] = expression("N"[D])
pretty[16] = expression("PNUE")
pretty[2] = expression("Asat"[mass])
pretty[7] = expression("N"[resorp])
pretty[12] = expression("Chl"[a:b])
pretty[15] = expression('WUE')
pretty[13] = expression('CC'[mass])
pretty[14] = expression('CC'[area])

#plot PCA biplot
pdf("Figure4.pdf",height=8,width=8)
par(mar=c(5,5,1,1),mfrow=c(1,1))
plot(pcaMethods::scores(ppca1)[,1],pcaMethods::scores(ppca1)[,2],xlab="PC1 (22%)",ylab="PC2 (18%)",cex.lab=1.5,cex.axis=1.4,,col=c(c1,c2,c3)[as.numeric(as.factor(dat$Region))], pch=19,cex=2,xlim=c(-7,7),ylim=c(-6,6))
points(pcaMethods::scores(ppca1)[,1],pcaMethods::scores(ppca1)[,2],xlab="PC1 (22%)",ylab="PC2 (18%)",cex.lab=1.5,cex.axis=1.4,,col=colvec[as.numeric(as.factor(dat$Region))], pch=21,cex=2,xlim=c(-7,7),ylim=c(-6,6),lwd=2)
f = 15
arrows(0,0,pcaMethods::loadings(ppca1)[,1]*f,pcaMethods::loadings(ppca1)[,2]*f,col=adjustcolor("gray50", alpha.f = .5),
       pch=21,cex=1,lwd=1.5, length=.07, angle=30)
for (i in 1:length(pcaMethods::loadings(ppca1)[,2])){
text(
  pcaMethods::loadings(ppca1)[i,1]*f*1.05,pcaMethods::loadings(ppca1)[i,2]*f*1.05,
      #pcaMethods::loadings(ppca1)[i,1]*f+if(-pcaMethods::loadings(ppca1)[i,1]>0){.2}else{-.2},
      #pcaMethods::loadings(ppca1)[i,2]*f+if(-pcaMethods::loadings(ppca1)[i,2]>0){.2}else{-.2},
     pretty[i],
     col="black",
     cex=.9,
     xpd=T, 
  )
}
#legend(legend=c("Japan","France","USA"),bty="n",cex=1.3,text.col=colvec[c(3,2,1)],"topleft")
text(-6,6,"Japan",col=colvec[3],cex=1.3)
text(-6,5.5,"France",col=colvec[2],cex=1.3)
text(-6,5,"USA",col=colvec[1],cex=1.3)
dev.off()

summary(aov(pcaMethods::scores(ppca1)[,1]~dat$Region))
TukeyHSD(aov(pcaMethods::scores(ppca1)[,1]~dat$Region))
  #Along PC1, ENA is different than both France and Japan

rev(sort(abs(pcaMethods::loadings(ppca1)[,1]))) #loadings sorted

summary(aov(pcaMethods::scores(ppca1)[,2]~dat$Region))
TukeyHSD(aov(pcaMethods::scores(ppca1)[,2]~dat$Region))
  #Along PC2, France is different than both Japan and ENA

rev(sort(abs(pcaMethods::loadings(ppca1)[,2]))) #loadings sorted

#save loadings
#write.csv(pcaMethods::loadings(ppca1),"loadings.csv")

#create summary statistic table (Table 1) of trait means in each region for pub

traitmat$Nmass = traitmat$Nmass*100
traitmat$leafC = traitmat$leafC*100
traitmat$resorbNeff = traitmat$resorbNeff*100
traitmat[,c(17:20,24:28)] = traitmat[,c(17:20,24:28)]*100


tabout = matrix(ncol=3,nrow=28,0) #output of means per region
tabset = matrix(ncol=3,nrow=28,0) #output of SEs per region
Fout = NULL #output of F stat per trait
Pout = NULL #output of P value for each F stat
tukout = matrix(ncol=3,nrow=28,0) #TukeyHSD P values per contrast
for(i in 1:dim(tabout)[1]) {
  tabout[i,] = tapply(traitmat[,i],fall$Region,function(x)mean(x,na.rm=T))
  tabset[i,] = tapply(traitmat[,i],fall$Region,function(x)(sd(x,na.rm=T)/sqrt(length(x[!is.na(x)]))))
  summary(aov(traitmat[,i]~fall$Region))
  Fout[i] = summary(aov(traitmat[,i]~fall$Region))[[1]][1,4]
  Pout[i] = summary(aov(traitmat[,i]~fall$Region))[[1]][1,5]
  #tukey stats
  tukout[i,] = TukeyHSD(aov(traitmat[,i]~fall$Region))[[1]][,4]
}

#tableout = data.frame(tabout,tabset,Fout,Pout,tukout)

#create text table
Jms = paste(signif(tabout[,1],3),signif(tabset[,1],2),sep=" \u00B1 ")
Fms = paste(signif(tabout[,2],3),signif(tabset[,2],2),sep=" \u00B1 ")
Ums = paste(signif(tabout[,3],3),signif(tabset[,3],2),sep=" \u00B1 ")
test = data.frame(Jms,Fms,Ums); rownames(test) = pretty
test

tableout = data.frame(test,round(Fout,2))
tableout
#write.table(tableout,"tableout.txt")

#tukey stats
data.frame(rownames(test),round(tukout,3))

```

Thus two 'shift syndromes' in away range:
1) ENA: more total N (and less resorption and PNUE), more Rubisco, more fiber but less CCmass, possibly minor defense
2) France: higher SLA, more light harvesting and Rubisco, less defense, lower photosynthetic capacity

Home Japanese populations maintain high photosynthetic rates and high WUE. 






